Customizing Software by Direct Manipulation of Tabular Data

### 2020-05-22 paper.tex
finish section 4

We have successfully used Wildcard to build customizations for 11
different websites. In Section~\ref{sec:evaluation}, we present
reflections from this process, outlining the kinds of customizations we
were able to build, limitations we encountered, and the challenges of
integrating scraping logic with websites.

@@

This scenario has shown a few examples of how table-driven
customizations can help a user improve their experience of a website.
Section~\ref{sec:evaluation} explains many other use cases, but first we
explain how the system works internally.

@@

In Wildcard, we have built three specific types of table adapters so
far.

@@

We have designed the table adapter API to be general enough to support
other types of useful adapters in the future. Here are two such
possibilities:

@@

In addition to the basic table editing operations, we also provide
\emph{cell editors}: UI widgets that expose a custom editing UI for a
single cell of the table view. A programmer building a cell editor need
only integrate it with the table viewer; propagating values into the
website UI is handled by the site-specific DOM adapter. In
Section~\ref{sec:evaluation} we provide some examples of using cell
editors.

@@

\hypertarget{sec:evaluation}{%
\section{Evaluation}\label{sec:evaluation}}

@@

We have used Wildcard to customize 11 websites, including transactional
sites like Amazon and Uber Eats, and media consumption sites like Hacker
News and Youtube. Table 1 summarizes these customizations, showing the
number of lines of code in the adapter configuration for each site.

@@

We have found that table-driven customization can serve a broad range of
useful purposes. Here we expand on some archetypal examples that
illuminate various aspects of using the system in practice.

@@

Sometimes, websites have opaque algorithms which they use to rank
results, which presumably maximize profit but don't offer much control
to the user. For example, Airbnb doesn't allow users to sort listings by
price, and Youtube doesn't allow users to sort the homepage by watch
time. Wildcard enables users to take back some control over this
process.

In other cases we have found that a lack of sorting options seems less
intentional, and simply represents a missing feature. For example, the
Instacart grocery delivery service doesn't allow users to sort their
grocery cart by price or category. Our guess is that they simply haven't
gotten around to implementing the feature yet.

In the current implementation of Wildcard, users can only sort and
filter entries that are shown on the current page, which means that
users are not entirely liberated from the suggestions of the opaque
algorithm. We could work around this restriction in the future by
fetching content across multiple pages, but we've also found that
sorting/filtering a single page of a paginated list is often an
acceptable outcome (and sometimes even a preferable one). For example,
it's more useful to sort (by length, say) 30 recommended Youtube videos
than to try to sort all videos on Youtube, or to sort (by read time,
say) the articles on the Hacker News front page than to sort all
non-archived articles.

@@

Many other web annotation systems focus on annotating text or arbitrary
webpage content, but Wildcard limits annotations to structured objects
extracted by an adapter, resulting in quite a different set of use
cases.

Annotating with Wildcard has proven most useful when taking notes on a
list of possible options (e.g., evaluating possible Airbnb locations to
rent). We have also used it with Instacart's online grocery cart, for
jotting down notes like ``should we get more milk?''

@@

Formulas are the most powerful part of the Wildcard system. So far, our
language supports only a small number of predefined functions. Adding
more will allow a very broad range of useful computations, as shown by
the success of spreadsheets.

Formulas are especially useful for fetching data from Web APIs. We've
used them to augment Airbnb listings with walkability scores, and to
augment Hacker News articles with estimated read times. One challenge of
the current language design is that supporting a new web API requires
writing conventional code to add a new function to the language, because
web APIs typically return complex JSON data structures that can't be
easily displayed in a single table cell. Adding functions for handling
JSON data to the formula language might make it possible to access new
APIs without requiring an entire new function in the formula system.

We have also found instances where simple data manipulation is useful:
adding together different subtotals into a total price, or manipulating
the results of a formula expression with arithmetic or concatenating
string labels, as shown in the example in Section~\ref{sec:example}.
(\emph{todo: this makes sense iff we switch the examples sectino to HN})

@@

An important benefit of cell editors is that they allow users to
incorporate their personal information within the web UI but without
uploading their personal data to the web server. To explore this idea,
we created a cell editor based on the FullCalendar Javascript plugin,
which can load data from a Google Calendar. This makes it convenient to
enter dates into a website based on the user's personal calendar
information.

Another benefit of cell editors is that a user can choose their
preferred widget for editing some type of information. We built a cell
editor based on the CKEditor rich text editor, and demonstrated its use
by integrating it with Google's Blogger website and using it to edit
blog posts, in place of Blogger's built-in editor.

@@

There are many useful customizations that are not currently possible in
the table-driven customization paradigm. Wildcard can only make
customizations that use the available data exposed in the table. UI
modifications are limited in scope; deleting arbitrary buttons isn't
possible, for example. At one point, we wanted to build an automation to
repeatedly load a grocery delivery website to check for open delivery
slots, but it wasn't clear how to achieve this. Some of these
limitations are specific to the current implementation of the Wildcard
extension, but some are more fundamental to the entire idea of
table-driven customization.

We consider this an acceptable outcome; our goal is to make many useful
customizations available to end users with a low threshold, and not to
span all possible customizations. Section 5.1 discusses this point
further.

One benefit of Wildcard is predictability: once we built a site adapter
for a website, it was generally obvious to us what types of
customizations were and weren't possible, and the UI guided us towards
building customizations that matched the system's model. If data is not
available in the table, for example, then it's abundantly clear that
customizations relying on that data will not be expressible.

\hypertarget{viability-of-dom-scraping}{%
\subsection{Viability of DOM scraping}\label{viability-of-dom-scraping}}

@@

Most of our DOM scraping adapters were created by members of our team.
However, an external developer unaffiliated with the project contributed
one adapter, designed to sort the Github page listing a user's
repositories. They described the experience as ``very straightforward.''

The adapters for our test sites ranged from 36 to 117 lines of code,
averaging 68 lines; Table 1 shows the full lines of code for each
adapter.

Some of the challenges of writing a DOM scraping adapter are the same
ones as with writing normal web scraping code, but the more interactive
nature of Wildcard introduces additional challenges. rve changes, but it
may prove challenging to observe changes on some sites only through the
DOM. One challenge is triggering updates to the spreadsheet data in
response to UI changes that happen after initial page load. Site
adapters are responsible for recognizing these changes by observing the
DOM. So far, we have been able to use event listeners and the
MutationObserver API to successfully obse

Another challenge is persisting updates to the DOM---some websites use
virtual DOM frameworks that can occasionally overwrite changes made by
Wildcard.

So far, we've managed to work around these issues for all the websites
we've tried, but we don't claim that any website can be customized
through DOM scraping. As web frontend code gets increasingly complex
(and starts to move beyond the DOM to other technologies like Shadow DOM
or even WebGL for rendering), it may become increasingly difficult to
customize websites from the outside without first-party support.

\emph{todo: mention ajax scraping} \emph{todo: mention privacy benefits}

@@

It is important to note that there are many specific customizations that
can be achieved in systems like Chickenfoot \citep{bolin2005} that
cannot be reproduced in Wildcard. We consider this an acceptable
tradeoff in exchange for a gentler slope in customization, and we show
in Section~\ref{sec:example} and Section~\ref{sec:evaluation} that our
model can still implement many useful customizations in practice. Also,
there is sometimes a way to reframe an imperative script in terms of our
direct manipulation model; for example, a script that iterates through
rows in a page adding some additional information can be reproduced
using a formula in Wildcard.

@@

>>>>

We have successfully used Wildcard to build customizations for 11
different websites. In Section~\ref{sec:reflections}, we present
reflections from this process, outlining the kinds of customizations we
were able to build, limitations we encountered, and the challenges of
integrating scraping logic with websites.

@@

This scenario has shown a few examples of how table-driven
customizations can help a user improve their experience of a website.
Section~\ref{sec:reflections} explains many other use cases, but first
we explain how the system works internally.

@@

We have built three specific types of table adapters so far, to enable
web customization using Wildcard.

@@

We have designed the table adapter API to be general enough to support
other types of useful adapters in the future. Here are two examples:

@@

In addition to the basic table editing operations, we also provide
\emph{cell editors}: UI widgets that expose a custom editing UI for a
single cell of the table view. A programmer building a cell editor need
only integrate it with the table viewer; propagating values into the
website UI is handled by the site-specific DOM adapter. In
Section~\ref{sec:reflections} we provide some examples of using cell
editors.

@@

\hypertarget{sec:reflections}{%
\section{Reflections on Usage}\label{sec:reflections}}

@@

We built Wildcard integrations for 11 websites, including transactional
sites like Amazon and Uber Eats, and media consumption sites like Hacker
News and Youtube. Table 1 summarizes these integrations, showing the
number of lines of code in the adapter configuration for each site, and
some example customizations we were able to perform for each site.

@@

We have found that table-driven customization can serve a broad range of
useful purposes. Here we expand on some archetypal examples that
illuminate aspects of using the system in practice.

@@

Sometimes, websites have opaque ranking algorithms which presumably
maximize profit but don't offer much control to the user. For example,
Airbnb previously allowed users to sort listings by price, but removed
that feature in 2012. We have found other sites where a lack of sorting
options seems more like merely an omission; for example, the Instacart
grocery delivery service has a spartan UI for viewing a pending order,
which doesn't allow for sorting by price or category. In both of these
cases, Wildcard enables users to take back some control.

In the current implementation of Wildcard, users can only sort and
filter entries that are shown on the current page, which means that
users are not entirely liberated from the suggestions of the opaque
algorithm. This restriction could be overcome in the future by scraping
content across multiple pages, or by using an integrated adapter built
in to the application. However, we've also found that sorting/filtering
a single page of a paginated list is often an acceptable outcome (and
sometimes even a preferable one). It's more useful, for example, to sort
30 recommended Youtube videos than to try to sort all videos on Youtube.

@@

Many web annotation systems focus on annotating text or arbitrary
webpage content, but Wildcard limits annotations to structured objects
extracted by an adapter, resulting in a different set of use cases.
Annotating with Wildcard has proven most useful when taking notes on a
list of possible options (e.g., evaluating possible Airbnb locations to
rent). We have also used it with Instacart's online grocery cart, for
jotting down notes like ``should we get more milk?''

@@

Formulas are the most powerful part of the Wildcard system. So far, our
language supports only a small number of predefined functions. Adding
more should allow a broad range of useful computations, as shown by the
success of spreadsheets.

Formulas are especially useful for fetching data from Web APIs. We've
used them to augment Airbnb listings with walkability scores, and to
augment Hacker News articles with estimated read times. One challenge of
the current language design is that supporting a new web API requires
writing Javascript code to add a new function to the language, because
web APIs typically return complex JSON data structures that can't be
easily displayed in a single table cell. In the future we would like to
make it possible to call new APIs without adding a dedicated function,
which might require adding functions to the formula language that can
manipulate JSON data.

We have also found instances where simple data manipulation is useful,
e.g.~transforming the results of an API call with basic arithmetic and
string operations, as shown in Section~\ref{sec:example}.

@@

We developed two cell editors to explore different benefits of adding
custom UI for editing the table.

An important benefit of cell editors is that they allow users to
incorporate their private information into a web UI without uploading it
to a website. To explore this idea, we created a datepicker based on the
\href{https://fullcalendar.io/}{FullCalendar} plugin, which can load
data from a Google Calendar. This makes it convenient to enter dates
into a website based on the user's personal calendar information.

Another benefit of cell editors is that a user can choose their
preferred widget for editing some type of information. We built a text
editor based on the \href{https://ckeditor.com/}{CKEditor} rich text
editor, and demonstrated its use by integrating it with Google's Blogger
website and using it to edit blog post content.

@@

There are many customizations that are not possible to implement with
table-driven customization. Some of the limitations are specific to the
current implementation of the Wildcard extension, but many are more
fundamental to the general paradigm.

One limitation is that Wildcard can only make customizations that use
the available data exposed in the table. If the adapter doesn't expose
some piece of data, the user can't use it in their customization. The
table data format also rules out customizing certain sites that don't
have a way to map to a table. The UI modifications available in Wildcard
are also limited in scope; deleting arbitrary buttons isn't possible,
for example. There is no facility for running automations when the user
isn't actively viewing a page---at one point, we wanted to build an
automation to repeatedly load a grocery delivery website to check for
open delivery slots, but it didn't seem possible to achieve this in
Wildcard.

Despite these limitations, we found that Wildcard does offer
predictability: once we built a site adapter for a website, it was
generally obvious what customizations were possible, and the UI guided
us towards building customizations that matched the system's model. If
data is not available in the table, for example, then it's abundantly
clear that customizations relying on that data will not be expressible.

We consider all of these limitations to be an acceptable outcome. Our
goal is to support as many useful customizations as possible with a low
threshold of difficulty, and not to span all possible customizations. If
users want to implement more sophisticated customizations, they have the
option of graduating to more advanced customization tools.

\hypertarget{viability-of-scraping}{%
\subsection{Viability of scraping}\label{viability-of-scraping}}

@@

Nearly all of our DOM scraping adapters were created by members of our
team. However, an external developer unaffiliated with the project
contributed one adapter, designed to sort the Github page listing a
user's repositories, and they described the experience as ``very
straightforward.''

The adapters for our test sites ranged from 36 to 117 lines of code,
averaging 68 lines; Table 1 shows the number of lines of code for each
adapter. Most of the code in the adapters is simply using DOM APIs and
CSS selectors to implement conventional web scraping logic.

Some of the challenges of writing a DOM scraping adapter are the same
ones as with writing normal web scraping code, but the more interactive
nature of Wildcard introduces additional challenges. One challenge is
registering appropriate event handlers to update the table data in
response to UI changes that happen after initial page load. Another
challenge is persisting updates to the DOM---some websites use virtual
DOM frameworks that can occasionally overwrite changes made by Wildcard.
So far, in practice we've managed to work around these issues for all of
the websites we've tried, but we don't claim that any website can be
customized through DOM scraping. As web frontend code gets increasingly
complex (and starts to move beyond the DOM to other technologies like
Shadow DOM or even WebGL), it may become increasingly difficult to
customize websites from the outside without first-party support.

AJAX scraping proved highly useful in several cases. The Uber Eats
website was challenging to scrape because it has a complex DOM structure
with machine-generated CSS classes, but the site also generates AJAX
requests which contain all the relevant data in a structured form that
is much easier to extract. We also found examples where relevant
information wasn't present in the DOM at all. On the grocery delivery
site Instacart, we used AJAX scraping to augment grocery items with
their categorization in the store, enabling use cases like viewing all
the vegetables in an order together.

@@

It is important to note that there are many specific customizations that
can be achieved in systems like Chickenfoot \citep{bolin2005} that
cannot be reproduced in Wildcard. We consider this an acceptable
tradeoff in exchange for a gentler slope in customization, and we show
in Section~\ref{sec:example} and Section~\ref{sec:reflections} that our
model can still implement many useful customizations in practice. Also,
there is sometimes a way to reframe an imperative script in terms of our
direct manipulation model; for example, a script that iterates through
rows in a page adding some additional information can be reproduced
using a formula in Wildcard.

@@

\emph{other platforms?}
